---
title: GRAPHICS WITH R - A CROWDSOURCE BASED TRAINING INITIATIVE
author:
  - name: Rudradev Sengupta
    affiliation: Johnson & Johnson
    address: Turnhoutseweg 30, 2340 Beerse, Belgium
    email: rsengup4@its.jnj.com
  - name: Paulo R. Bargo
    affiliation: Johnson & Johnson
    address: Spring House, PA 19477, United States
    email: pbargo@its.jnj.com
  - name: Bill Pikounis
    affiliation: Johnson & Johnson
    address: Spring House, PA 19477, United States
    email: bpikounis@its.jnj.com
  - name: Surya Mohanty
    affiliation: Johnson & Johnson
    address: Spring House, PA 19477, United States
    email: smohanty@its.jnj.com
  - name: Jason Milnes
    affiliation: RStudio, PBC
    address: Pittsburgh, PA, United States
    email: Jason@rstudio.com
  - name: Philip Bowsher
    affiliation: RStudio, PBC
    address: Pittsburgh, PA, United States
    email: Phil@rstudio.com
abstract: >
  Over the past decades, SAS has been a powerful tool in the data analysis repertoire of pharmaceutical statisticians. However, the recent development of automation capabilities based on R such as RMarkdown and R/Shiny have created new channels to speed up creation of reports, presentations and interactive graphics. Moreover, R helps to standardize these processes for different phases of a drug development or submission process \citep{Brodsky2012}. At the Johnson & Johnson subsidiary Janssen, we aim to improve the collective literacy in R programming across the entire statistician community and achieve nearly 100% adherence by pharmaceutical statisticians and statistical programmers over the next couple of years. In order to achieve this goal, we are currently leveraging all types of training formats, from online training, to in-house instructor-led seminars, to one-on-one mentoring. Using RStudio Cloud, as a platform for internal crowd-led hands-on workshops in particular, has been an integral part of training statisticians/programmers to solve on-the-job problems and business needs ranging from visualization to automated reports. In this paper, we discuss our journey of crowdsource-based training.
output:
  rticles::rjournal_article:
    includes:
      in_header: preamble.tex
---

## Introduction

<!-- Introductory section which may include references in parentheses \citep{R}, or cite a reference such as \citet{R} in the text. -->

<!-- For many years SAS has been a powerful tool in the data analysis repertoire of pharmaceutical statisticians however the recent development of automation capabilities such as RMarkdown and R/Shiny have created a new venue to expedite access to consumable information in the form of reports, presentations or interactive graphics that can be produced efficiently and in standard format for all phases of a drug development or submission process. As the Pharmaceutical sector boosts its interactions with regulatory agencies using R programming as one key instrument for drug development submissions, we face a dilemma that several members of statistics and statistical programming teams are not currently advanced R programmers. At Janssen we aim to improve the literacy in R programming and achieve nearly 100% adhesion by statistics and statistical programming teams in the coming 2-3 years. To achieve this goal, we are leveraging all types of training formats, from online training, to in-house instructor led seminar, to one-on-one mentoring. One effective method we have been developing is the use of RStudio.Cloud as a platform for internal crowdsourced hands-on workshops where statisticians/programmers are “thought” to solve on-the-job real problems ranging from visualization to automated reports.  -->
Crowdsourcing is a methodology that is becoming widely used to solve different problems (Howe 2006). Perhaps one of the best examples of crowdsourcing in practice is Wikipedia (https://en.wikipedia.org/wiki/Wikipedia). In education, crowdsourcing has been used to create content \citep{Hills2015}, assessment material development \citep{Alghamdi2015}, and practical experience for learners \citep{Chen2014}, amongst other jobs. 
In this paper, we will discuss our experience of a crowdsource-based program led by our statisticians for the training of Janssen personnel in visualization with R, using solely internal and open source resources. The ``Graphics with R" initiative was crafted with the intention of increasing scientific communication through graphical presentations, specifically through R. Most members of clinical statistics and statistical programming teams were not well versed in R programming, a skill set that is becoming essential in the current landscape. The need to acquire this new skill and the lack of time for large cohorts of team members to engage in regular “in-person” classes led us to a crowdsource-based approach where new “on-the-job” content was created by more experienced volunteers from the “crowd”. The same volunteers supported each other and trained less experienced team members in a “co-learning/co-teaching” web that successfully helped us train 120 novice R programmers over a span of 8 months in a workflow that required 2-4 hours of training every other week. 

\begin{figure}[h]
  \centering
  \scalebox{0.3}{\includegraphics{figures/crowdsourceFramework3.png}}
  \caption{Crowdsourcing framework.}
  \label{figure_csfrmwrk}
\end{figure}

## Our approach

Figure \ref{figure_csfrmwrk} asks four basic questions to setup a crowdsource-based training framework: 

* What is the incentive?
* Who is the crowd?
* What to crowdsource?
* How to crowdsource?

We developed the plan for our training keeping these questions in mind. Each of them are addressed in the next sections to explain in details the bi-weekly approach we took to achieve our goal. 

### What is the incentive?

Crowdsourcing can be a powerful alternative method for training, bringing in some additional benefits. One of the key aspects of crowdsource-based training is the need for all participants to seek knowledge as well as share it, in a co-learn/co-teach framework. Participants with previous knowledge in R were integral part of this initiative for the development of the curricula as well as to lead the group meetings and become the central point of contact. An active network was naturally created within smaller groups of people who were already familiar with each other, which greatly facilitated the exchange of knowledge, forming a support-group for all participants. This also helped the leaders of the program to delegate some of their jobs and avoid getting bombarded with questions and requests all the time on top of their regular project work. This was particularly important in a corporate setting because no leaders were professional educators and had to maintain a balance between their regular jobs and the work done as part of this training. In that sense crowdsourcing was implemented in several levels: creation of content, hands on training, review of performance, and creation/presentation of advanced content (tips and tricks), with everyone contributing. 

Another benefit of the process implemented was that the entire workload was distributed over a large span of time which required a smaller amount of work per week and a continuous engagement. Therefore the participants had to maintain engagement and repeat their efforts in a regular manner, which helped retaining the learning. Beyond the learning of R, this format also helped the participants to develop soft skills as all participants had opportunities to showcase their learning in presentations during the regroup meetings as well as opportunities to develop some leadership skills when leading the groups.

### Who is the crowd?

At first, a survey was done to know the proportion of statisticians with a background in R or who were already using R on a day-to-day basis for regular projects. As Figure~\ref{figure_Ruser} displays, among the respondents identified as non-clinical statisticians more than 80\% of the colleagues already had a background in R. In contrast, for clinical statistics group, this proportion was close to one third. Hence, the main focus audience for the training were clinical statisticians.

\begin{figure}[htbp]
  \centering
  \scalebox{0.8}{\includegraphics{figures/Ruser.png}}
  \caption{Proportion of statisticians trained in R.}
  \label{figure_Ruser}
\end{figure}

All the participants were divided into different workspaces according to their therapeutic areas (TAs) or region they are located in. This method of crowdsource-based training helped us to also divide the entire workload among a group of experts who were already experienced and regular users of R. We managed to identify approximately 30 volunteers to help us with this initiative in different roles as described below (refer to Figure \ref{figure_vlntrgroups}):

* Leads:
  + Administrate platform.
  + Lead regroup sessions: US, Non-US (duplicate sessions to take care of the participants from different time zones across the world).
* Group Leaders:
  + Lead assignment by region/TA.
  + Lead status review meeting.
  + Define presenters for regroup meeting.
  + Help create assignments.
* Assistants:
  + Help create assignments.
  + Lead individual troubleshooting sessions.
  + Present tips and tricks for visualization with R.


\begin{figure}[htbp]
  \centering
  \scalebox{0.45}{\includegraphics{figures/vlntrgroups2.pdf}}
  \caption{Division of available volunteers into different groups.}
  \label{figure_vlntrgroups}
\end{figure}

### What to crowdsource?

The main objective of this initiative was to train statisticians to create informative graphics by using recently developed R packages, e.g., `ggplot2`. There were several challenges to plan this training, e.g. the different levels of expertise within existing community of statisticians, difficulty to set aside time for training, difficulty to retain knowledge with standard methods of training, etc. These were all key items which we had to take into account before planning this initiative. Finally, it was decided to start from simple exercises and then keep building on top of this foundation to arrive at more complex assignments such as animated plots, interactive plots, etc., by the end of the course. The detailed syllabus is listed below:

* Assignment 1: Univariate Plots.
* Assignment 2: Univariate Plots by Grouping Variable.
* Assignment 3: Bivariate Plots.
* Assignment 4: Linear regression.
* Assignment 5: Longitudinal data analysis.
* Assignment 6: Survival analysis - Kaplan-Meier, Hazard Functions, etc. 
* Assignment 7: ROC Curves.
* Assignment 8: Forest Plots.
* Assignment 9-10: RMarkdown.
* Assignment 11-12: Arranging Data for Clear Communication.
* Assignment 13: Tidyverse (tidyr and dplyr) and By Processing.
* Assignment 14: Heatmaps.
* Assignment 15: Interactive and Animated plots.
* Assignment 16: Interactive Dashboards with Flexdashboard and Shiny.

\noindent
**Examples**

The above syllabus was created in consultation with our team of experts and represented typical graphics used in drug submission packages from the different TAs. Data had to be blinded or simulated as RStudio Cloud access did not exist within Janssen firewall. Every assignment consisted of a dataset together with a data description file and questions related to that particular assignment. After a few weeks, feedback was collected from all the participants, both learners and internal expert instructors. While initially, the first couple of assignments were prescriptive, in desired results, it was decided to add hints and make some questions of the assignment optional based on the feedback from the TA leads. On one hand, it managed the workload for the beginners to gain skills and confidence. On the other hand, interested participants could further explore with their intellectual curiosity. The materials for all the assignments are available [online](https://github.com/rsengup4/GraphicsWithR). 

\begin{figure}[htbp]
  \centering
  \scalebox{0.38}{\includegraphics{figures/picCollage.pdf}}
  \caption{Showcasing different examples from the course.}
  \label{figure:picCollage}
\end{figure}

As a specific example, let us discuss the first assignment and its possible solutions. Many assignments were kept open to interpretation of the participants as we wanted to motivate everyone to come up with their own solutions, rather than trying to follow one particular approach to solve the assignment. 

<!-- $\vspace{2cm}$ -->
\noindent
**Assignment 1**

\noindent
**Data**

This data is taken from the R package `TH.data` \citep{Hothorn}. For the data frame object GBSG2, there were 686 observations of the following 10 variables:

* horTh: Hormonal therapy, a factor at two levels no and yes
* age: Age of the patients in years
* menostat: Menopausal status, a factor at two levels pre (premenopausal) and post (postmenopausal).
* tsize: Tumor size (in mm)
* tgrade: Tumor grade, a ordered factor at levels I < II < III.
* pnodes: Number of positive nodes.
* progrec: Progesterone receptor (in fmol).
* estrec: Estrogen receptor (in fmol).
* time: Recurrence free survival time (in days)
* cens: Censoring indicator (0- censored, 1- event)

\noindent
**Question**

Associated data was provided in an Excel file (Assign1.xlsx) and participants were asked to solve the following questions:

* Produce a scatterplot for progesterone receptor (in fmol) vs. estrogen receptor (in fmol). Add color and shape to include information on
  + Tumor Grade
  + Menopausal Status
* Produce pairwise plots of progesterone receptor (in fmol) vs. estrogen receptor (in fmol) vs. tumor size. It should include additional information through color and shape to provide better understanding of data.

\noindent
**Solution**

As the variable distributions were skewed, new variables were created using the log transformation. There were also 0 values present which required a basic scoring rule where 0 values were replaced by 1.

Solutions from different teams generally fell along the following 2 ways: 

* Plot original data but the scale will be log-scale (Figure 5).

* Plot the log-transformed data (Figure 6).

* `ggpairs()` function was used to create the pairwise plot of the three variables as specified in the last part of the assignment (Figure 7). 

```{r hw1plot1, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Scatterplot for progesterone receptor (in fmol) vs. estrogen receptor (in fmol) in log-scale. Tumor grade and menopausal status are differentiated by the color and shape, respectively.", out.width = '80%', fig.pos="h",fig.align="center"}
## Loading the required libraries
library(readxl)
library(ggplot2)
library(GGally)

## Reading the data
HW1_data <- read_excel("data/Assign1.xlsx")
HW1_data$progrec2 <- ifelse(HW1_data$progrec==0, 1, HW1_data$progrec)
HW1_data$estrec2 <- ifelse(HW1_data$estrec==0, 1, HW1_data$estrec)

## Plotting Original Data on log-scale
fig1a<- ggplot(data= HW1_data, aes(x=progrec2, y=estrec2, col=tgrade, shape=menostat)) 
fig1a + geom_point() + 
        labs(x="Progesterone receptor (fmol)", y="Estrogen receptor (fmol)", 
             title="Effect of Progesterone Receptor on Estrogen Receptor", 
             subtitle="Scatterplot of Estrogen and Progesterone", 
             caption="Data source: Assign1.xlsx \n 
             Original data on log scale (0 was imputed to 1)", col="Tumor Grade",
             shape="Menopausal Status") + 
        scale_x_log10() + 
        scale_y_log10() +
        theme_classic() + 
        theme(plot.title=element_text(color="red4", size=10, face="bold", 
             lineheight=0.9, hjust=0.5), plot.subtitle=element_text(color="red4", 
             hjust=0.5,size=9), plot.caption=element_text(color="gray", face="italic", 
             size=8)) 

```


```{r hw1plot2, echo=FALSE, fig.cap="Scatterplot for log-transformed progesterone receptor (in fmol) vs. log-transformed estrogen receptor (in fmol). Tumor grade and menopausal status are differentiated by the color and shape, respectively.", out.width = '80%',fig.pos="h",fig.align="center"}

## Plotting log-transformed data - use the log function within the aes() function
fig1b <- ggplot(data= HW1_data, aes(x=log10(progrec2), y=log10(estrec2), col=tgrade, 
       shape=menostat)) 

fig1b + geom_point() + 
        labs(x="Progesterone receptor (fmol)", y="Estrogen receptor (fmol)", 
            title="Effect of Progesterone Receptor on Estrogen Receptor", 
            subtitle="Scatterplot of Estrogen and Progesterone", 
            caption="Data source: Assign1.xlsx \n 
            Log transformed data (0 was imputed to 1)", col="Tumor Grade", 
            shape="Menopausal Status") +
        theme_classic() + 
        theme(plot.title=element_text(color="red4", size=10, face="bold", 
             lineheight=0.9, hjust=0.5), plot.subtitle=element_text(color="red4", 
             hjust=0.5,size=9), plot.caption=element_text(color="gray", face="italic", 
             size=8)) 
```


```{r hw1plot3, echo=FALSE, fig.cap="Pairwise Scatterplot of Progesterone, Estrogen and Tumor Size. Tumor grade and menopausal status are differentiated by the color and shape, respectively.", out.width = '90%', fig.pos="h",fig.align="center"}

HW1_data$progrec3 <- log10(HW1_data$progrec2)
HW1_data$estrec3 <- log10(HW1_data$estrec2)

fig2 <- ggpairs(data=HW1_data, columns=c("progrec3", "estrec3", "tsize"), 
                title="Pairwise Scatterplot of Progesterone, Estrogen and Tumor Size",
                axisLabels="show", 
                columnLabels=c("Progesterone (fmol)","Estrogen (fmol)","Tumor Size"), 
                upper=list(continuous="points"), 
                mapping=aes(color=tgrade, shape=menostat), legend=c(3,2), 
                diag="blank", 
                xlab=NULL, 
                ylab=NULL) 
fig2 +  theme_bw() + 
        theme(plot.title=element_text(color="red4", size=10, face="bold", 
              lineheight=0.9, hjust=0.5), 
              plot.subtitle=element_text(color="red4", hjust=0.5,size=9), 
              plot.caption=element_text(color="gray", face="italic", size=8)) +
        labs(caption="Data source: Assign1.xlsx \n 
              Log transformed data for progesterone and estrogen (0 imputed to 1)", 
              col="Tumor Grade", shape="Menopausal Status")
```


### How to crowdsource?

RStudio Cloud was used as the primary platform during the entire course of the training. In the beginning, we encountered some challenges with the initial set up as we were using the Beta version of the RStudio Cloud platform (the production environment was not available at that time). RStudio team was prompt to resolve these issues by opening some of the features such as the number of workspaces and projects per account. As a starting point, we set up different workspaces for different TAs and RStudio engineers helped us configure those workspaces properly. The `assignment` feature of RStudio Cloud helped us to easily distribute the assignments for all the workspaces. Different features available for RStudio Cloud are discussed in the next subsection.

\vspace{0.35cm}

\noindent
**About RStudio Cloud**

RStudio Cloud is a hosted version of RStudio in the cloud. It is maintained by RStudio and has a nice web interface together with certain features which made it the most suitable option for us. 

\begin{figure}[htbp]
  \centering
  \scalebox{0.23}{\includegraphics{figures/RStudioCloudOverview}}
  \caption{Overview of RStudio Cloud.}
  \label{figure:rscloud}
\end{figure}

Many organizations are embarking on data science educational initiatives and academies to help bring employees up to speed on state of the art analytical topics. RStudio Cloud is a hosted version of RStudio that continues the mission and dedication to a sustainable investment in free and open-source software for data science, scientific research, education, and technical communication. RStudio Cloud makes it easy for anyone to do, share, teach, and learn data science, by allowing students and instructors to:
 
* Analyze data using the RStudio IDE directly from a browser
* Share projects and analysis with teams, classrooms, workshops, or the world
* Use interactive tutorials and primers to learn the basics of data science
* No hardware, configuration, or installation required

One of the most effective ways to get started learning R is to start using it. RStudio Cloud is designed to make educational data science efforts easier for instructors and students by removing complicated setup and environment configuration. Projects are the fundamental unit of work on RStudio Cloud - projects encapsulate R code, packages and data files and provide isolation from other analyses or course work. Projects can be public or private and are created and managed from a user’s workspace. Public projects can allow people to interact with an analysis by visiting a shared link. Every RStudio Cloud user gets a personal workspace in which to create projects. Users can also create private, shared spaces that function as virtual classrooms for courses and workshops. Users who can access a specific workspace can be assigned different roles, giving them capabilities appropriate for instructors, teaching assistants and students.

When an instructor is teaching a class or workshop, RStudio projects can be made into assignments and academic content for students. Students can work on the assignments and instructors can assess their work and check progress. Students can also make copies of class projects created by the instructor, with the necessary environment and packages automatically replicated. Instructors can access usage metrics for any shared space where they are an admin or moderator. They can view summary information for the space, and data for individual members and their projects. Summary metrics include total project hours and the number of active spaces, members, and projects. These integrations mean that instructors do not need to spend time managing environments or troubleshooting student technical issues. Instructors can focus on what is most important - teaching data science. 

\noindent
**Setting up RStudio Cloud**

Six separate workspaces were set up for five different TAs and one specifically for the China region. The "Assignment" feature of RStudio Cloud was used to post the assignments in every workspace from GitHub. This feature helped the participants to easily create their own copies of the assignments and then work on it on their own time. Initially, the "Base Project" feature was used to control the package management so that the beginners can directly start working on the assignments without worrying about R package installation. Later participants were encouraged to install the packages on their own to get more familiar with R

<!-- \vspace{0.25cm} -->
\noindent
**Assignment workflow**

At week 0, the first assignment was posted in all the six workspaces in RStudio Cloud and all the participants were notified so they could start working on the assignments as soon as they were ready. Then the group leaders organized TA status review sessions during week 1 to discuss the assignment within the corresponding groups. One colleague from each TA would be selected to present the solution on behalf of the group, in the regroup session happening at week 2. The status review sessions also helped the group members that stayed behind, either due to difficulties understanding/performing the task or due to lack of time because of daily projects, to catch up. During the regroup sessions in week 2 all groups presented their solutions, and discussion on the different approaches helped all participants learn more about other tools and approaches within R. One "tips and tricks" presentation, with examples on R-based graphics in everyday work, was hosted during these regroup sessions to showcase how R can help us visualize and explore pharmaceutical data as well as showcase advanced R programming techniques. Also, during the regroup sessions some hints and background information for the next assignment were discussed. After every regroup session the recorded videos and the solutions by all groups were made available through standard sharing platforms. The next assignment was than made available in RStudio Cloud within 24 hours and everyone was notified. This iterative biweekly process was used for all 16 assignments over 8 months. Besides, the regular workflow as displayed in Figure \ref{figure_workflow}, different subgroups hosted their within-group sessions as per requirements.

\begin{figure}[htbp]
  \centering
  \scalebox{0.35}{\includegraphics{figures/biweekly.pdf}}
  \caption{Bi-weekly approach for the assignments.}
  \label{figure_workflow}
\end{figure}

## Project metrics
The training started with a virtual kick-off meeting where the objectives and the format of the course was laid out. More than 200 people attended the kick-off meeting as it was a group-wide initiative. The first assignment had 170 participants. However, that participation quickly dropped to between 100 to 120 participants per week by the time the third assignment was handed out and this number held steady until the end of the training. 

A second survey was sent to the participants at the conclusion of the training and was completed by about half (N=56) of the cohort that participated until the end. About 40% of the respondents had never used R and after the training 36% felt they managed to achieve a beginners' level proficiency in R. Of that 36% indicating proficiency, 50% were at intermediate level and 13% at advanced level. 63% of the respondents said they participated on all or almost all the sessions. A little over 50% thought the assignments were challenging. All but 2 respondents thought the training was a valuable use of their time and 3 said they would not like to continue learning R. One third of the respondents said they are now using R in their day-to-day activities. We also questioned how much time they spent on average every week to complete the assignments. 43% spent at most a couple of hours, whereas 34% had to work for 2-4 hours, and 13% worked for more than 4 hours every week to finish the assignments.

## Management perspective

The Graphics with R initiative was generated from a high-level goal of the head and senior leadership team of the entire statistician organization within the Janssen R & D subsidiary of Johnson & Johnson. The umbrella strategic goal was centered on leveraging new technologies to build our capabilities to add value in alignment with mission components of driving robust and scientifically informed decision-making to discover and develop medicines for patients in need. Raising the organization's collective competency in graphics, using the software environment R, was settled upon as an achievable and concrete target. A charter was developed. A quantitative metric was set to reach at least 75% participation and proficiency across more than 200 clinical statisticians in the community. Messaging was clear from the very top leadership of the statistics organization that creating graphics with R was a required objective for everyone.

It was especially gratifying how broadly the statistics community took this charter and its principles. The entire community was willing to iterate and continuously get better individually via crowdsourcing. They obtained the necessary skills, predominantly through (1) self learning with online resources and (2) working with colleagues in a collaborative manner. There were multiple cases of colleagues, from early through later stages of their careers, who had no prior knowledge of R, yet surprised themselves in learning how to efficiently develop and successfully display visualizations, often within the 2 weeks period between assignments described earlier. This had substantial side benefits for raising other collective competencies of reporting, data wrangling, and coding across the organization. 

By the second assignment presentation, the core oversight team decided that `ggplot` would be required as the sole system within R to fulfill the remaining assignments. Similarly, by the fourth assignment presentation, R Markdown or the RStudio IDE was required by the core team for all participants. These midstream iterations were generally well-received and followed by participants. The core team had communicated at the start of the series that such iterations were likely as the crowdsourcing framework was being used for the first time within the organization.

One high performer even took the opportunity to apply their new graphics and R skills to an ongoing real-time health authority dialogue.  A visualization helped persuade agreement on use of a biomarker for more accurate prediction of survival in cancer patients.  

Ongoing next steps are to broaden the use of graphics with R and R Markdown for rapid delivery of topline reports (TLR). TLRs are important channels for top executive reviews and decisions of the truly critical statistical results after the completion of clinical trials.

## Summary and discussion

RStudio Cloud was an excellent tool to introduce the participants to R and manage the distribution of the exercises as well as capture essential metrics on the engagement. We did not need to worry about R installations or access permissions to internal cloud resources as RStudio Cloud already took care of these details inherently. Also, by distributing the exercises through the RStudio Cloud project we could seamlessly include all dependencies, such as packages, with the necessary instructions to complete the work.

As the program was designed with great flexibility there were important lessons, learnt throughout the journey, which were continuously used as feedback to improve the program. The set up of the exercises without the obligation of producing a specific outcome provided different teams an opportunity to create diverse solutions on how to achieve the end results. For instance, the initial focus of the training was the creation of graphics using R but as the training developed some groups introduced the use of RMarkdown to complete the exercise, and after exposing the entire group to this new tool and the use of one of the tips and tricks sessions to delve into the technology, all the groups were able to implement this new aspect of the training. Another examples of this diversity was the introduction of R packages beyond those installed in RStudio Cloud for the completion of the exercises. The same flexibility caused some discomfort to some of the participants, particularly in the beginning of the training. However, we managed to overcome this challenge via the tips and tricks sessions as well as by providing hints for certain questions of the assignments. It is also important to highlight that although we maintained great flexibility throughout the course, adapting as we moved along, the project started with a great deal of planning where discussions with all volunteers helped craft the proper curriculum that catered to the diverse needs of all TAs, as different graphics are commonly used in their daily work. Moreover, there was a huge background effort to identify and define the best platform (RStudio Cloud) to be used during the course for programming and sharing the information.

Although strongly encouraged by upper management this training was non-mandatory therefore we believe that overall the training was successful as a participation rate between 60%-70% (from initial dates) was maintained. Besides the strong participation another key metric for success was the fact that a third of the users was using what they learnt on their daily jobs after the conclusion of the training. Part of this success can be attributed to the design of the training as participants engaged on the assignments every week for an extended period of time therefore becoming more comfortable with the use of the language. This certainly helped with retaining knowledge.

From the management perspective the Graphics with R initiative proved to be an efficient method to train a large group of colleagues with minimal disruption on the daily operations of the group. The crowdsourcing format was crucial as learners took ownership on their learning and dedicated themselves at their own pace. Beyond the actual advancement of programming skills the opportunity to develop soft skills was also greatly appreciated by the management team and was evident in raising the broad collective capabilities on communication skills. Due to this success two other training initiatives adapted crowdsourcing methods and were implemented. Therefore we believe crowdsourcing is a powerful method for modern training in the professional setting.  

## Acknowledgement

We would like to acknowledge all the volunteers who agreed to help us out during this journey in various means e.g., by creating the assignments, by presenting during tips and tricks sessions, by helping their colleagues with one-one sessions to help them improve, etc. Janssen colleagues who were invaluable to this initiative: Katleen Callewaert, Yannick Vandendijck, Beatriz Lopez Sanchez, Etienne Ernault, Monelle Tamegnon, Lixia Pei, Cynthia Gargano, Jianfeng Xu, Bin Gao, Xiang Li, Iva Kezic, Xin Qiu, Sabrina El Bachiri, Yan Liu, Xiaoming Li, Jinxiang Wu, Weiping Liu, Stefan Avey, Volha Tryputsen, Martin Otava, Davit Sargsyan, Ewoud de Troyer, Jocelyn Sendecki, Oluyemi Oyeniran, Marjolein Crabbe, Leacky Muchene, Jan Serroyen, Jeroen Tolboom, Leen Slaets, Sumesh Kalappurakal and Axel Muehlig. We also want to thank Robby Shaver from RStudio for his invaluable support with the RStudio Cloud platform.

<!-- This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf). -->

\bibliography{RJreferences}
